---
title: "Reproducible Research knitr assignment"
author: "KellyPisane"
date: "June 15, 2017"
output: html_document
---


First, unzip and read in the data and look at the head of the data frame. 
```{r readdata}
unzip("activity.zip")
activity <- read.csv("activity.csv")
head(activity)
```

## Question: What is the mean total number of steps taken per day?

Calculate the total number of steps taken per day
```{r dailysteps}
dailyTotals <- as.data.frame(tapply(activity$steps, activity$date, sum, na.rm = TRUE))
colnames(dailyTotals) <- c("totalSteps")
head(dailyTotals)
```

Make a histogram of the total number of steps taken each day
```{r histogram}
dates <- rownames(dailyTotals)
rownames(dailyTotals) <- NULL
dailyTotals <- cbind(dates, dailyTotals)
colnames(dailyTotals) <- c("dates", "totalSteps")
hist(dailyTotals$totalSteps, xlab = "Total Steps (daily)", ylab = "Frequency", main = "Daily Steps", col = "red")

```
Make a histogram of the total number of steps taken each day
Calculate and report the mean and median of the total number of steps taken per day