---
title: "Reproducible Research knitr assignment"
author: "KellyPisane"
date: "June 15, 2017"
output: html_document
---
```{r options, echo=FALSE}
options(scipen = 999)
```

First, unzip and read in the data and look at the head of the data frame. 
```{r readdata}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## Question: What is the mean total number of steps taken per day?  

###Calculate the total number of steps taken per day  
```{r dailysteps}
dailyTotals <- as.data.frame(tapply(activity$steps, activity$date, sum, na.rm = TRUE))
colnames(dailyTotals) <- c("totalSteps")
head(dailyTotals)
```

###Histogram of the total number of steps taken each day  
```{r histogram}
dates <- rownames(dailyTotals)
rownames(dailyTotals) <- NULL
dailyTotals <- cbind(dates, dailyTotals)
colnames(dailyTotals) <- c("dates", "totalSteps")
hist(dailyTotals$totalSteps, xlab = "Total Steps (daily)", ylab = "Frequency", main = "Daily Steps", col = "red")
```

###The mean and median of the total number of steps taken per day are:    
```{r meanmedian}
medianTotal <- median(dailyTotals$totalSteps)
meanTotal <- mean(dailyTotals$totalSteps)
```

mean = `r meanTotal`  
median = `r medianTotal`  

##What is the average daily activity pattern?  

```{r averageInterval}
intervalMeans <- as.data.frame(tapply(activity$steps, activity$interval, mean, na.rm = TRUE))
intervals <- as.numeric(rownames(intervalMeans))
rownames(intervalMeans) <- NULL
intervalMeans <- cbind(intervals, intervalMeans)
colnames(intervalMeans) <- c("intervals", "meanSteps")
head(intervalMeans)
```

### Plot the mean number of steps taken in each time interval across all days  

```{r plotMeanSteps}
with(intervalMeans, plot(intervals, meanSteps, type = "l", main = "Mean Steps per 5 Minute Interval", xlab = "Time (minutes)", ylab = "Mean Number of Steps"))

```

### Determine which interval has the highest mean number of steps

```{r maxSteps}
i <- which(intervalMeans$meanSteps == max(intervalMeans$meanSteps))
highest <- intervalMeans$intervals[i]
highValue <- intervalMeans$meanSteps[i]
```
The highest mean number of steps is `r highValue` and occurs at interval `r highest`.  

## Inpute missing values  

### Total number of missing values 

```{r missingValues}
NAs <- sum(is.na(activity$steps))
```
  
There are `r NAs` NAs in the dataset.  

To fill in the missing values, use the mean number of steps in that interval. 

```{r fillIn}
activity2 <- activity
for(i in 1:17568){
    if(is.na(activity2$steps[i])){
        activity2$steps[i] <- intervalMeans$meanSteps[which(intervalMeans$intervals == activity2$interval[i])]
    }
}
```
  
The dataset activity2 is the same as the original dataset but with the missing values filled in.  

```{r show}
head(activity2)
```

### Recalculate the daily total number of steps from the imputed data.  

```{r dailysteps2}
dailyTotals2 <- as.data.frame(tapply(activity2$steps, activity2$date, sum, na.rm = TRUE))
colnames(dailyTotals2) <- c("totalSteps")
head(dailyTotals2)
```

### Re-plot the histogram of the daily total number of steps including the imputed data.  

```{r histogram2}
dates <- rownames(dailyTotals2)
rownames(dailyTotals2) <- NULL
dailyTotals2 <- cbind(dates, dailyTotals2)
colnames(dailyTotals2) <- c("dates", "totalSteps")
hist(dailyTotals2$totalSteps, xlab = "Total Steps (daily)", ylab = "Frequency", main = "Daily Steps", col = "red")
```

###The mean and median of the total number of steps taken per day calculated from the imputed data are:    
```{r meanmedian2}
medianTotal2 <- median(dailyTotals2$totalSteps)
meanTotal2 <- mean(dailyTotals2$totalSteps)
```

mean = `r meanTotal2`  
median = `r medianTotal2`  


###Effects of imputing data.  

The mean total number of steps increases as does the median total number of steps.  
The biggest difference can be seen in the shape of the two histograms. There are many more occurences that fall into the center of the distribution.  The reason for this is that every point that has been added to the dataset was an average value and so falls near the center of the distribution.  
