---
title: "Reproducible Research knitr assignment"
author: "KellyPisane"
date: "June 15, 2017"
output: html_document
---


First, unzip and read in the data and look at the head of the data frame. 
```{r readdata}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## Question: What is the mean total number of steps taken per day?  

###Calculate the total number of steps taken per day  
```{r dailysteps}
dailyTotals <- as.data.frame(tapply(activity$steps, activity$date, sum, na.rm = TRUE))
colnames(dailyTotals) <- c("totalSteps")
head(dailyTotals)
```

###Histogram of the total number of steps taken each day  
```{r histogram}
dates <- rownames(dailyTotals)
rownames(dailyTotals) <- NULL
dailyTotals <- cbind(dates, dailyTotals)
colnames(dailyTotals) <- c("dates", "totalSteps")
hist(dailyTotals$totalSteps, xlab = "Total Steps (daily)", ylab = "Frequency", main = "Daily Steps", col = "red")
```

###The mean and median of the total number of steps taken per day are:    
```{r meanmedian}
medianTotal <- median(dailyTotals$totalSteps)
meanTotal <- mean(dailyTotals$totalSteps)
```

mean = `r meanTotal`  
median = `r medianTotal`  

##What is the average daily activity pattern?  

```{r averageInterval}
intervalMeans <- as.data.frame(tapply(activity$steps, activity$interval, mean, na.rm = TRUE))
intervals <- as.numeric(rownames(intervalMeans))
rownames(intervalMeans) <- NULL
intervalMeans <- cbind(intervals, intervalMeans)
colnames(intervalMeans) <- c("intervals", "meanSteps")
head(intervalMeans)
```

### Plot the mean number of steps taken in each time interval across all days  

```{r plotMeanSteps}
with(intervalMeans, plot(intervals, meanSteps, type = "l", main = "Mean Steps per 5 Minute Interval", xlab = "Time (minutes)", ylab = "Mean Number of Steps"))

```

### Determine which interval has the highest mean number of steps

```{r maxSteps}
highest <- which(intervalMeans$meanSteps == max(intervalMeans$meanSteps))
highest <- intervalMeans$intervals[highest]
```
The highest mean number of steps occurs at interval(s) of `r highest`.  